---
title: "Reactivity tutorial"
output: html_document
runtime: shiny
---

*Reactive expressions*


Load the shiny library, if needed
```{r}
library(shiny)
```

This is a working app inspired by and adapted from https://shiny.posit.co/r/articles/build/action-buttons/

The app generates n random values (where n is a user input) and creates a histogram of the values.

```{r}
library(shiny)

ui <- fluidPage(
  numericInput("n", "n", 50),
  plotOutput("plot"),
)

server <- function(input, output) {
  
  output$plot <- renderPlot({
    randomVals <- runif(input$n)
    hist(randomVals)
  })
}

shinyApp(ui, server)
```

You decide you want to add a second plot of the random values, this time a boxplot. But there's an issue - the random values for plots 1 and 2 are a different set of randomly generated values. Use a reactive expression to generate the random values once to be used for both plot 1 and plot 2.

```{r}
library(shiny)

ui <- fluidPage(
  numericInput("n", "n", 50),
  plotOutput("plot1"),
  plotOutput("plot2")
)

server <- function(input, output) {
  
  output$plot1 <- renderPlot({
    randomVals <- runif(input$n)
    hist(randomVals)
  })
  
  output$plot2 <- renderPlot({
    randomVals <- runif(input$n)
    boxplot(randomVals)
  })
}

shinyApp(ui, server)
```

*Action buttons*

You have another app (it looks strangely like the last one) but this time it is very slow. (This is because of the Sys.sleep line that makes R wait for 3 seconds every time the plot updates. I've used it to simulate code that is slow to compute.)

```{r}
library(shiny)

ui <- fluidPage(
  numericInput("n", "n", 50),
  plotOutput("plot"),
)

server <- function(input, output) {
  
  output$plot <- renderPlot({
    Sys.sleep(3) # You can move this line but do not delete it
    randomVals <- runif(input$n)
    hist(randomVals)
  })
}

shinyApp(ui, server)
```

Edit this app to have an action button, so that the plot only updates when the user clicks the button.

```{r}
library(shiny)

ui <- fluidPage(
  numericInput("n", "n", 50),
  plotOutput("plot"),
)

server <- function(input, output) {
  
  output$plot <- renderPlot({
    Sys.sleep(3) # You can move this line but do not delete it
    randomVals <- runif(input$n)
    hist(randomVals)
  })
}

shinyApp(ui, server)
```
